---
title: "Example Application"
output: html_notebook
---

# Introduction

This notebook shows how to use the `rampdata` package in an application.
This package will add several features to the main function, if we
initialize it to use those features.

1. Logging that will be saved specially on the cluster.
2. Provenance of files saved with each file and provenance of
   the main function saved in the provenance directory beside
   the data.
3. All paths specified relative to the data directory that's
   configured for each user and machine. This lets you run from
   the cluster or from home.
4. Input files and output files can be specified to have a
   particular version by giving a workflow configuration file.


# Main

## The main function

We put code into packages, and one or more functions in that package
are top-level functions. A researcher or a workflow tool will call
those functions directly. These are main functions, and the `rampdata`
package helps write them so that they run better on the cluster.

A main function can either be called interactively, at an R command line,
or from the command line as `R -e "function()"`, or from the command line
as a script. In the last case, you write a script, such as
```{r sample_script, eval = FALSE}
gisdemog::create_uganda_cover()
```
I would save that single line into a file `uganda_cover.R` and run it
from the command line with `Rscript uganda_cover.R --shapefile uganda.shp`.

## The usual layout of a main function

A main function function usually separates sanity-checking tasks
and input/output tasks from the work of the function. For something
meant to run from the command line, it might look as follows.
```{r old_main, eval = FALSE}
create_uganda_cover <- function(shapefile, landscape) {
  inputs <- read_input_files(shapefile, landscape)
  outputs <- calculate_uganda_cover(inputs$shapefile, inputs$lanscape)
  write_outputs(outputs$cover)
}
create_uganda_cover_main <- do.call(create_uganda_cover, cover_parse_args(commandArgs()))
```
The first function is ready to call from the R command line.
The second function is ready to call from the Bash command line
on the cluster.

## An updated version of the main function

```{r updated_main, eval = FALSE}
create_uganda_cover <- function(shapefile, landscape) {
  rampdata::prov_begin()
  tryCatch({
    inputs <- read_input_files(shapefile, landscape)
    outputs <- calculate_uganda_cover(inputs$shapefile, inputs$lanscape)
    write_outputs(outputs$cover)
  }, finally = {
    rampdata::prov_end()
  }
}
create_uganda_cover_main <- function() {
  do.call(create_uganda_cover, cover_parse_args(commandArgs()))
}
```


# Argument parsing and initialization

When we parse arguments to the code, we need to initialize

1. Logging - Tell it whether we are on the cluster or
   on a local machine. When on the cluster, warnings, errors,
   and fatal messages will be saved to a separate file.
   We can also change the logging level for the whole
   application or for particular packages or R scripts.
   
2. Workflow versions - This is a file with lists of versions
   and the file paths they affect.

The location of the data directory is initialized as needed
from a configuration file in the home directory.

```{r options_parsing}
.cover.doc <- 'Create Uganda Coverage

Usage:
  create_uganda_coverage <workflow> [-v] [-q] [--local]

Options:
  -v       Set logging to debug level.
  -q       Set logging to warn level.
  --local  Working on a local machine, not the cluster.
'
cover_parse_args <- function(args = NULL) {
  if (is.null(args)) {
    args <- commandArgs(TRUE)
  }
  parsed <- docopt::docopt(.cover.doc, args = args, version = "1.0")
  
  calling_function <- deparse(sys.call(sys.parent(n = 1)))
  set_logging_from_args(
    packageName(), calling_function, parsed$v, parsed$q, parsed$local
    )
  initialize_workflow(args$workflow)
}
```


