---
title: "Path Interface"
output: html_notebook
---

## Introduction
I need to make an interface to set a path for RAMP. The first goal of the
path handling is to give our files a common structure that we
recognize for searching and file management. The second goal is to make
that structure have a root directory that we can change on the cluster
or can set for local compuation on desktops.


## Alternatives

The first version of an interface asked the user to pass in a list.
```{r first_version, eval = FALSE}
dest <- local_path(
  list(stage = "output", project = "uganda2020", dataset = "district_age2")
)
```

That seems like using named arguments, but with extra steps.

The options are the following. Use paths.
```{r paths_version, eval = FALSE}
dest <- local_path("/uganda2020/output/district_age2/1")
```
Use arguments.
```{r args_version, eval = FALSE}
dest <- local_path(
  stage = "output", project = "uganda2020", dataset = "district_age2", version = "1"
)
```
Use paths with replacement.
```{r replace_version, eval = FALSE}
dest <- local_path("/uganda2020/output/district_age2/$version")
```

## Scenarios
After using this path for a while, these are some things that I noted
I wanted to do.

Give me that file.
```{r that_file, eval = FALSE}
filename <- local_path("/uganda2020/output/district_age2/1/pop_2000.csv")
```

Give me a root directory to ensure it exists.
```{r rooted, eval = FALSE}
dest <- local_path("/uganda2020/output/district_age2/1")
if (!dir.exists(dest)) dir.create(dest)
filename <- file.path(dest, paste0("pop_", year, ".csv"))
```

The Python FBDPath we used for the Forecasting Burden of Disease
group had a nice feature that would look like this:
```{r add_path, eval = FALSE}
dest <- local_path(project = "uganda2020")
dest <- add_path(dest, stage = "mic", version = "3")
dest <- add_path(dest, dataset = "density", file = "file_2000.csv")
```
That was added by people who used the parsed path a lot, so it's
likely a helpful feature. The trouble is that there
isn't a strong conversion capability in R. We could make a sensible
conversion.
```{r as_path, eval = FALSE}
read.csv(as.path(dest))
```

